---
title       : "Shiny App: Calculating BMI"
subtitle    : A slidify presentation for the Data Products course
author      : Fabian van Beveren
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## About

To show what we learned in this coursera course, this Shiny app was created. 
It calculates the BMI, based on the input:
- Height (cm)
- Weight (kg)

---

## Calculations

The simplest part is the calculation of the BMI itself.

```{r}
#The function to calculate the BMI
calcbmi <- function(height, weight) round(weight/(height/100)^2, 2)

#Example
height <- 168
weight <- 60
calcbmi(height, weight)
```

---

## Calculations

Furthermore, to determine whether the person is of normal weight or not, the limits 
for the given height are calculated.

```{r}
#Calculation
calclimits <- function(height) {
  i <- numeric(3)
  i[1] <- 18.5 * (height/100)^2 #<18.5 = underweight
  i[2] <- 25 * (height/100)^2   #18.5 - 25 = normal weight
  i[3] <- 30 * (height/100)^2   #25 - 30 = overweight | >30 = obese
  i
}

#Example
calclimits(168)
```

---

## Plot

A plot is created to show how close you are to a healthy/unhealthy weight. Code is 
not shown in the slidify presentation, but can be found in the Rmd and of course in 
the Shiny server files.

```{r echo=FALSE, fig.width=10,fig.align='center'}
#plot function
plotbmi <- function(height, weight, limits) {
  y <- calcbmi(height, x)
  plottitle <- paste("The BMI for a height of", as.character(height), "cm, for a given weight")
  plot(x, y, 
       main = plottitle, 
       xlab = "Weight", ylab = "BMI", 
       type = "l", lty = 5)
  #The black line is the input weight
  abline(v = weight, lwd = 2, col = "navy")
  #The following lines are weights that correspond to the limit values, to give some
  #visual aid. Makes it easier to determine your category visually.
  abline(v = limits[1], lty = 2, lwd = 1.5, col = "yellow") 
  abline(v = limits[2], lty = 2, lwd = 1.5, col = "orange")    
  abline(v = limits[3], lty = 2, lwd = 1.5, col = "red")
  #Add text in the plot
  text(x = c(17, (limits + 1)), 
       y = c(max(y)*.85, max(y)*.89, max(y)*.95, max(y)), 
       adj = c(0, 1), 
       labels = c("Underweight", "Normal weight", "Overweight", "Obese"))
}
lims <- calclimits(168)
x <- 1:21*5 + 15
plotbmi(168, 60, lims)
```